Predicting the (non-cryptographic) PCG Pseudo-Random Number Generator In Practice

### 2020-05-31 article/main.tex
cas retenue

	\begin{align*}
	\lvert 2^{k-\ell}\widetilde{T}_i-U_i \rvert &=\lvert 2^{k-\ell}(T_i + N_i)-U_i \rvert \\
	 &= \lvert 2^{k-\ell}N_i - U_i[0,k-\ell] \rvert\\
	 &\leqslant \lvert 2^{k-\ell}N_i\rvert + \lvert U_i[0,k-\ell] \rvert \\
	 &\leqslant 2^{k-\ell} + 2^{k-\ell}-1\\
	 &\leqslant 2^{k-\ell +1 }- 1
	\end{align*}
	Hence \(\lVert U -2^{k-\ell} \widetilde{T} \rVert \leqslant \sqrt{n}(2^{k-\ell+1}-1) \)

	\item If there are indexes such that \(\widetilde{T}_i = 2^\ell-1\).
	\begin{itemize}
		\item If \(N_i \in \{0,1\}\), then \(\widetilde{T}_i = T_i + N_i\) (without modulo) and \[\lvert 2^{k-\ell}\widetilde{T}_i-U_i \rvert \leqslant \leqslant 2^{k-\ell +1 }- 1\]
		\item If \(T_i = 0\) and \(N_i = -1\), then \(U_i=U_i[0,k-\ell]\) and  \[\lvert 2^{k-\ell}\widetilde{T}_i-(U_i + 2^k) \rvert =\lvert 2^{k-\ell}(2^\ell -1) -(U_i + 2^k) \rvert = \lvert 2^{k-\ell} + U_i[0,k-\ell]\rvert \leqslant 2^{k-\ell + 1} -1 \]
	\end{itemize}
	We need a new vector \(U'\) defined as follow.
	\begin{itemize}
		\item If \(T_i = 0\) and \(N_i = -1\), \(U'_i=U_i+2^k\).
		\item Else \(U'_i=U_i\).
	\end{itemize}
	In both case \(\lVert U' -2^{k-\ell} \widetilde{T} \rVert \leqslant \sqrt{n}(2^{k-\ell+1}-1) \)
	\item If there are indexes such that \(\widetilde{T}_i = 0\).
	\begin{itemize}
		\item  If  \(N_i \in \{-1,0\}\), then \(\widetilde{T}_i = T_i + N_i\) (without modulo) and \[\lvert 2^{k-\ell}\widetilde{T}_i-U_i \rvert \leqslant \leqslant 2^{k-\ell +1 }- 1\]
		\item If \(N_i = 1\) we have a problem.
	\end{itemize}
	So, if there are indexes such that \(\widetilde{T}_i = 0\), we have to guess if \(N_i=1\) or not.

\end{itemize}

@@

In this article, we will only use an exact CVP solver when \(k = 128\) and \(\ell = 6\). As discussed in the previous subsection, we suppose we are never in the situation\(\widetilde{T} = 2^\ell-1\) and \(N_i=1\). Hence \(\lVert T-U \rVert \leqslant \sqrt{n}(2^{k-\ell+1}-1) \). So, if we fix \(\ell = 6\) we can search for the minimum \(n\) such that \[2\sqrt{n}2^{122} \leqslant \lambda_1(L(128,n))\].

After computation, we find that the minimal \(n\) is 42. \todo{a vérifier (snif snif)}

@@

\begin{table}[htbp!] \todo{a vérifier}
  \centering
  \begin{tabular}{|c|c|c|c|}
	\hline
	n & \(\lVert G(n,64) \rVert \times \lVert G(n,64)^{-1} \rVert\) & \( \lambda_1(L(n,64)) \) & minimum \(\ell\) \\
	\hline
	3 & 2.87 & \(4.09e^{12} \simeq 2^{41.9}\) & 19 \\
	4 & 2.06 & \(2.44e^{14} \simeq 2^{47.8}\) & 13 \\
	5 & 3.77 & \(1.72e^{15}\simeq 2^{50.6}\) & 11 \\
	6 & 2.69 & \(1.03e^{16}\simeq 2^{53.2}\) & 8 \\
	\hline
  \end{tabular}
  \caption{minimal \(\ell\) needed for a given \(n\)}
  \label{tab:babai}
\end{table}

>>>>

	\begin{align*}
	\lvert 2^{k-\ell}\widetilde{T}_i-U_i \rvert &=\lvert 2^{k-\ell}(T_i + N_i)-U_i \rvert \\
	 &= \lvert 2^{k-\ell}N_i - U_i[0,k-\ell] \rvert\\
	 &\leqslant \lvert 2^{k-\ell}N_i\rvert + \lvert U_i[0,k-\ell] \rvert \\
	 &\leqslant 2^{k-\ell} + 2^{k-\ell}-1\\
	 &\leqslant 2^{k-\ell +1 }- 1
	\end{align*}

	\item If there are indexes such that \(\widetilde{T}_i = 2^\ell-1\).
	\begin{itemize}
		\item If \(N_i \in \{0,1\}\), then \(\widetilde{T}_i = T_i + N_i\) (without modulo) and \[\lvert 2^{k-\ell}\widetilde{T}_i-U_i \rvert \leqslant \leqslant 2^{k-\ell +1 }- 1\]
		\item If \(N_i = -1\), then \(U_i=U_i[0,k-\ell]\) and  \[\lvert 2^{k-\ell}\widetilde{T}_i-(U_i + 2^k) \rvert =\lvert 2^{k-\ell}(2^\ell -1) -(U_i + 2^k) \rvert = \lvert 2^{k-\ell} + U_i[0,k-\ell]\rvert \leqslant 2^{k-\ell + 1} -1 \]
	\end{itemize}

	\item If there are indexes such that \(\widetilde{T}_i = 0\).
	\begin{itemize}
		\item  If  \(N_i \in \{-1,0\}\), then \(\widetilde{T}_i = T_i + N_i\) (without modulo) and \[\lvert 2^{k-\ell}\widetilde{T}_i-U_i \rvert \leqslant \leqslant 2^{k-\ell +1 }- 1\]
		\item If \(N_i = 1\) , then \(U_i= 2^k-2^(k-l) + U_i[0,k-\ell]\) and  \[\lvert 2^{k-\ell}\widetilde{T}_i-(U_i - 2^k) \rvert =\lvert 0 -(U_i - 2^k) \rvert = \lvert 2^k-2^(k-l) + U_i[0,k-\ell] - 2^k\rvert \leqslant 2^{k-\ell + 1} -1 \]
	\end{itemize}
\end{itemize}
	We need a new vector \(U'\) defined as follow.
\begin{itemize}
	\item If \(T_i = 0\) and \(N_i = -1\), \(U'_i=U_i+2^k\).
	\item If \(T_i = 2^\ell - 1\) and \(N_i = +1\), \(U'_i=U_i-2^k\).
	\item Else \(U'_i=U_i\).
\end{itemize}
In all this cases \(\lVert U' -2^{k-\ell} \widetilde{T} \rVert \leqslant \sqrt{n}(2^{k-\ell+1}-1) \), and \(U' \equiv U \mod 2^k \).

\todo{C'est un lemme mais pour l'instant y'a pas de commande lemme}
\begin{theorem}
	If $\widetilde{T}$ is the noisy truncation of \(U\) of \(\ell\) bits, then there exists \(U'\) such that :
	\begin{itemize}
		\item \(U'\equiv U \mod 2^k\)
		\item \(\lVert U' -2^{k-\ell} \widetilde{T} \rVert \leqslant \sqrt{n}(2^{k-\ell+1}-1) \)
	\end{itemize}
\end{theorem}

@@

In this article, we will only use an exact CVP solver when \(k = 128\) and \(\ell = 6\). So, if we fix \(\ell = 6\) we can search for the minimum \(n\) such that \[2\sqrt{n}(2^{128-6+1}-1) \leqslant \lambda_1(L(128,n))\].

After computation, we find that the minimal \(n\) is 63.

@@

\begin{table}[htbp!]
  \centering
  \begin{tabular}{|c|c|c|c|}
	\hline
	n & \(\lVert G(n,64) \rVert \times \lVert G(n,64)^{-1} \rVert\) & \( \lambda_1(L(n,64)) \) & minimum \(\ell\) \\
	\hline
	3 & 2.87 & \(4.09e^{12} \simeq 2^{41.9}\) & 26 \\
	4 & 2.06 & \(2.44e^{14} \simeq 2^{47.8}\) & 20 \\
	5 & 3.77 & \(1.72e^{15}\simeq 2^{50.6}\) & 18 \\
	6 & 2.69 & \(1.03e^{16}\simeq 2^{53.2}\) & 15 \\
	\hline
  \end{tabular}
  \caption{minimal \(\ell\) needed for a given \(n\)}
  \label{tab:babai}
\end{table}
